---
title: "Southampton: Mapping Energy Poverty & Deprivation"
author: "Ben Anderson"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: TRUE
  bookdown::pdf_document2:
    toc: yes
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE) # set this to FALSE if you don't want the code in the final output

# useful R libraries
library(ggplot2)
library(readr)
library(dplyr)

# GIS libraries
library(raster)
library(sf)
library(spData)
library(spDataLarge)

dataFolder<-"/Users/megha/Desktop/1. Uni/Masters/Project/Mapping Code/data/" # default for Meghan
user <- Sys.info()[[7]] # who are we logged in as?
if(user == "ben"){
  dataFolder <- path.expand("~/University of Southampton/HCC Energy Landscape Mapping project - Documents/General/data/")
}

# this will not print anything out when knitted due to the include=FALSE setting in the chunk header
message("User: ", user)
message("dataFolder: ", dataFolder)
```

# Intro

Mapping fuel poverty, deprivation & doemstic energy use in Southampton to focus on areas which are high fp & high deprivation.

Based on:

 * [Meghan Kinglsey-Walsh's MSc](https://github.com/dataknut/energyMapping) project code
 * Tom Rushby's [open geography](https://dataknut.github.io/mapping-with-r/ONS-open-geography.html) example.

# Mapping Domestic Electricity consumption (LSOAs)

```{r Loading in the LSOA electricty consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/LSOA_ELEC_2019.csv")
library(data.table)
lsoa_elecData_dt <- data.table::as.data.table(readr::read_csv(inFile))
lsoa_elecData_dt$LSOA11CD <- lsoa_elecData_dt$`Lower Layer Super Output Area (LSOA) Code`

selectSoton <- function(dt){
  # function we can re-use
  # assumes a data.table
  # assumes LA name = la_name (may need to make new var first)
  select_dt <- dt[la_name %like% "Southampton"]
  return(select_dt)
}

lsoa_elecData_dt$la_name <- lsoa_elecData_dt$`Local Authority Name`
# filter Soton
lsoa_soton_elecData_dt <- selectSoton(lsoa_elecData_dt)

```

```{r Loading in boundary data for the LSOA}
#The LSOA boundaries for the Solent have been pre-downloaded
inf<-paste0(dataFolder, "boundaries/LSOA/lsoa_solent.shp")
message ("Loading LSOA boundaries from file")
lsoa_sf_data <- sf::read_sf(inf)
table(lsoa_sf_data$LAD11NM)#How many LSOAs are there in each LA
```

Focus on Southampton

```{r Mapping the electricty consumption of the LSOAs}
lsoa_merged_sf <- merge(lsoa_sf_data, lsoa_soton_elecData_dt) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf) + 
  geom_sf(aes(fill = `Mean domestic electricity consumption \n(kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all LSOAs)")
#Mapping it
```

# Mapping Domestic Gas (LSOAs)

```{r Loading in the LSOA gas consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/LSOA_GAS_2019.csv")
lsoa_gasData_dt <- as.data.table(readr::read_csv(inFile))
lsoa_gasData_dt$LSOA11CD <- lsoa_gasData_dt$`Lower Layer Super Output Area (LSOA) Code`
lsoa_gasData_dt$la_name <- lsoa_gasData_dt$`Local Authority Name`
# filter Soton
lsoa_soton_gas_dt <- selectSoton(lsoa_gasData_dt)
```

```{r Mapping the gas consumption of the LSOAs}
lsoa_merged_sf_gas <- merge(lsoa_sf_data, lsoa_soton_gas_dt) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf_gas) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Southampton (all LSOAs)")
#Mapping it
```

> Some areas are very high gas...

### Correlating gas & elecricity

```{r gaselecCorr}
setkey(lsoa_soton_elecData_dt, LSOA11CD)
lsoa_soton_elecData_dt[, meanEleckWh := `Mean domestic electricity consumption \n(kWh per meter)`]
setkey(lsoa_soton_gas_dt, LSOA11CD)
lsoa_soton_gas_dt[, meanGaskWh := `Mean consumption (kWh per meter)`]
dt <- lsoa_soton_elecData_dt[lsoa_soton_gas_dt]
ggplot2::ggplot(dt, aes(x = meanEleckWh, y = meanGaskWh)) +
  geom_point()
```


# Mapping the Indices of Multiple Deprivation (LSOAs)

2019 data

Should be a negative correlation with gas & electricity

```{r Loading in the Indices of Multiple Deprivation data}
# Multiple deprivation data at MSOA pre downloaded
inFile <-paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2019.csv")
lsoa_multidep_dt <- as.data.table(readr::read_csv (inFile))
lsoa_multidep_dt[, la_name := LADnm]
lsoa_multidepSoton_dt <- selectSoton(lsoa_multidep_dt)
```

```{r Mapping the Indices of deprivation for the LSOA}
lsoa_multidepSoton_dt$LSOA11CD <- lsoa_multidepSoton_dt$lsoa11cd #creating a variable with the LSOA code in the same name as in sf_data

lsoa_multidep_merged_sf <- merge(lsoa_sf_data, lsoa_multidepSoton_dt) #merging these
ggplot2::ggplot(lsoa_multidep_merged_sf) +
  geom_sf(aes(fill=`IMDScore`))+ # which data set has the scores?
  scale_fill_continuous(name="IMD Score", low="green",high="red")+
  labs(caption ="Southampton (all LSOAs)")
```

Check correlations with energy

```{r energyCorrIMD}
setkey(lsoa_multidepSoton_dt, LSOA11CD)
dt <- lsoa_soton_elecData_dt[lsoa_soton_gas_dt][lsoa_multidepSoton_dt]

plotDT <- dt[, .(meanEleckWh, meanGaskWh, IMDScore)]
pairs(plotDT)
```

> Much stronger relationship between mean gas use & IMD score in Southampton

# Mapping Fuel Poverty (LSOAs)

2019 Fuel Poverty
```{r Loading in the LSOA 2021 Fuel Poverty data}
inFile<-paste0(dataFolder, "Fuel Poverty/2019 LSOA_Fuel_Poverty.csv")
lsoa_fuelpov_2019_dt <- as.data.table(readr::read_csv (inFile))
lsoa_fuelpov_2019_dt[, la_name := `LA Name`]
lsoa_fuelpovSoton_2019_dt <- selectSoton(lsoa_fuelpov_2019_dt)
```

```{r Mapping the LSOA 2021 Fuel Poverty}
lsoa_fuelpovSoton_2019_dt[, pcFP := as.numeric(`Proportion of households fuel poor (%)`)]
lsoa_fuelpovSoton_2019_dt[, LSOA11CD := `LSOA Code`] #creating a variable with the LSOA code in the same name as in sf_data
lsoa_fuelpov_2021_merged <- merge(lsoa_sf_data, lsoa_fuelpovSoton_2019_dt) #merging these
# as above - why does it load as a char?

ggplot2::ggplot(lsoa_fuelpov_2021_merged) +
  geom_sf(aes(fill=pcFP))+ 
  scale_fill_continuous(name="Proportion of households fuel poor (%)", low="green",high="red")+
  labs(caption ="Southampton (all LSOAs)")
```

# Correlating energy use & fuel poverty/deprivation

Figure \@ref(fig:energyCorrAll) shows correlation of fuel poverty & IMD & energy

```{r energyCorrAll, fig.cap="Correlation of IMD score, % in fuel poverty and mean energy demand (Southampton LSOAs)"}
setkey(lsoa_fuelpovSoton_2019_dt, LSOA11CD)
dt <- lsoa_soton_elecData_dt[lsoa_soton_gas_dt][lsoa_multidepSoton_dt][lsoa_fuelpovSoton_2019_dt]

plotDT <- dt[, .(`MSOA Name`, LSOA11CD, 
                 meanEleckWh, meanGaskWh, IMDScore, IMDRank0, IMDDec0, IncScore, pcFP)]
pairs(plotDT[, .(meanEleckWh, meanGaskWh, IMDScore, IncScore, pcFP)])
```

IMD & % in fuel poverty correlate. Less clear correlation with mean energy. Need to check definitions... Also we might expect more of a correlation with the Income score (although INC score seems to drive IMD score in Southampton so...)

# Selecting areas with high fuel poverty & high deprivation

Create a local quantile for each - where are the places with higest deprivation & fuel poverty?

```{r checkAreas}
# reuse plotDT from above
# create local IMD & fuel poverty quintiles (5) 0->20% etc
plotDT[, IMD_quinSoton := cut(IMDScore, quantile(IMDScore))]
plotDT[, pcFP_quinSoton := cut(pcFP, quantile(pcFP))]

# how many LSOAs are in each cell of the crosstab
t <- plotDT[, .(n = .N), keyby = .(IMD_quinSoton, pcFP_quinSoton)]

t
```

Strange - why are there NAs?

Select them...

```{r selectAreas}

sample <- plotDT[pcFP_quinSoton  == "(12,25]" & 
              IMD_quinSoton == "(36.1,67.2]"][order(-pcFP)]
sample[, .(LSOA11CD, IMDScore, IMDDec0, IMD_quinSoton, pcFP, pcFP_quinSoton)]
message("Which national IMD decile are they in & what is mean % FP?")
sample[, .(mean_pcFP = mean(pcFP),
           mean_IMDScore = mean(IMDScore),
           nLSOAs = .N), keyby = .(IMDDec0)]
```
 
 Map them
 
```{r map areas}
map_df <- merge(lsoa_sf_data, sample) #merging the boundaries and energy data
ggplot2::ggplot(map_df) + 
  geom_sf(aes(fill = pcFP)) +
  scale_fill_continuous(name = "% fuel poverty", low = "green", high = "red") +
  labs(caption = "Southampton (selected LSOAs)")


ggplot2::ggplot(map_df) + 
  geom_sf(aes(fill = IMDScore)) +
  scale_fill_continuous(name = "IMD Score", low = "green", high = "red") +
  labs(caption = "Southampton (selected LSOAs)")

library(plotly)
p <- ggplot2::ggplot(sample, aes(x = IMDScore, y = pcFP)) +
  geom_point(aes(label = LSOA11CD))

plotly::ggplotly(p)
```

> Note that some areas are high IMD but relatively low % fuel poverty

Figure \@ref(fig:mapLeaflet) shows a leaflet version for pretty background to identify locations

```{r mapLeaflet, fig.cap="Location of filtered LSOAs"}
# https://dataknut.github.io/mapping-with-r/ONS-open-geography.html
library(leaflet)

st_coord_sys <- st_crs(map_df) # check coord system
st_coord_sys # current coord system EPSG: 4326 (is what leaflet wants - good)

# transform the coord system if required
if(st_coord_sys$epsg != 4326){
  map_df_trans <- st_transform(map_df, "+proj=longlat +datum=WGS84")
}

leaflet(map_df_trans) %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addPolygons(color = "blue", fillColor = "blue", fillOpacity = 0.2, weight = 1.5, popup = ~(LSOA11CD), # popups clicked
              highlight = highlightOptions(
                weight = 5,
                color = "#666",
                fillOpacity = 0.7,
                bringToFront = TRUE))

```

# The end

```{r environment}
R.Version()

```
