---
title: "Solent region: Test Maps"
author: "Meghan Kinglsey-Walsh"
date: 'Last run at: `r Sys.time()`'
output:
  bookdown::html_document2:
    fig_caption: yes
    code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 4
    toc_float: TRUE
  bookdown::pdf_document2:
    toc: yes
    fig_caption: yes
    number_sections: yes
  bookdown::word_document2: 
    fig_caption: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
    fig_width: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # set this to FALSE if you don't want the code in the final output

# useful R libraries
library(ggplot2)
library(readr)
library(dplyr)

# GIS libraries
library(raster)
library(sf)
library(spData)
library(spDataLarge)

dataFolder<-"/Users/megha/Desktop/1. Uni/Masters/Project/Mapping Code/data/" # default for Meghan
user <- Sys.info()[[7]] # who are we logged in as?
if(user == "ben"){
  dataFolder <- path.expand("~/University of Southampton/HCC Energy Landscape Mapping project - Documents/General/data/")
}

# this will not print anything out when knitted due to the include=FALSE setting in the chunk header
message("User: ", user)
message("dataFolder: ", dataFolder)
```


# Mapping the Local Authority

## Mapping Domestic Electricity Consumption

```{r loading in Local Authority electricity consumption data}
inFile <-paste0(dataFolder, "energy/Subnational_electricity_consumption_statistics_2019.csv")
la_elecData <- readr::read_csv(inFile)
summary(la_elecData)
```

```{r Loading in the Boundary Data for local authorities we are intrested in}
# geo_query <- "https://ons-inspire.esriuk.com/arcgis/rest/services/Administrative_Boundaries/Local_Authority_Districts_December_2018_Boundaries_UK_BGC/MapServer/0/query?where=lad18nm%20IN%20(%27Southampton%27,%27Portsmouth%27,%27Winchester%27,%27Eastleigh%27,%27Isle%20of%20Wight%27,%27Fareham%27,%27Gosport%27,%27Test%20Valley%27,%27East%20Hampshire%27,%27Havant%27,%27New%20Forest%27,%27Hart%27,%27Basingstoke%20and%20Deane%27)&outFields=lad18cd,lad18nm,long,lat&outSR=4326&f=geojson"
# #This only includes the ones we are interested in i.e the Solent
# message("Loading LA geometry from ONS Open Geography API") #Specifying the location message
# la_sf_data <- sf::st_read(geo_query)

# The above is very clever but it takes time
# Load from file
la_shpFile <- paste0(dataFolder, "/boundaries/LA/la_solent.shp")
la_sf_data <- sf::st_read(la_shpFile)
head(la_sf_data)
```

```{r Mapping the local authority electricty consumption}
la_elecData$lad18cd <- la_elecData$`LA Code`#Creating a variable with the local authority code in the same name of the consumption data
la_merged_sf <- merge(la_sf_data, la_elecData)#Merge the local Authority boundaries and the consumption data
ggplot2::ggplot(la_merged_sf) +
  geom_sf(aes(fill = Total_GWh))+scale_fill_gradient(low="green",high="red")

```

## Mapping Transport
#### Freight
```{r Loading the  Freight Transport consumption data for the Local Autority}
inFile <-paste0(dataFolder, "energy/Local Authority level Transport data.csv")
la_Transport <- readr::read_csv(inFile)
summary(la_Transport)
head(la_Transport)
```

```{r Mapping the local authority energy consumption (Freight)}

la_Transport$lad18cd <- la_Transport$Code # Creating a variable with the local authority code in the same name of the consumption data
la_merged_Transp <- merge(la_sf_data, la_Transport)# Merge the local Authority boundaries and the Transport data
ggplot2::ggplot(la_merged_Transp) +
  geom_sf(aes(fill = Freight)) +
  scale_fill_gradient(name = "Freight: Thousand tonnes of oil equivalent (ktoe)", low="green",high="red")
```

#### Personal

Note: this data is not per household/person so it will correlate (to some extent) with population size.

```{r Mapping the local authority energy consumption (Personal)}
# re-use the merged dataset
ggplot2::ggplot(la_merged_Transp) +
  geom_sf(aes(fill = Personal)) +
  scale_fill_gradient(name = "Personal: Thousand tonnes of oil equivalent (ktoe)", low="green",high="red") 
```
## Mapping Fuel Poverty 
### 2019 Proportion of Population

```{r Loading in the 2019 Local Authority Fuel Poverty Data}
inFile<-paste0(dataFolder, "Fuel Poverty/2019 LA_Fuel_Poverty.csv")
la_fuelpov_2019<-readr::read_csv(inFile)
head(la_fuelpov_2019)
```
```{r Mapping Local Authority Fuel Poverty 2019}
la_fuelpov_2019$lad18cd <- la_fuelpov_2019$`Area Code` # Creating a variable with the local authority code in the same name 
la_merged_Fuelpov2019 <- merge(la_sf_data, la_fuelpov_2019)# Merge the local Authority boundaries and the Fuel Poverty data
la_merged_Fuelpov2019$pcFP <- as.numeric(la_merged_Fuelpov2019$`Proportion of households fuel poor (%)`) # convert to numeric from char - why does it load as char??
ggplot2::ggplot(la_merged_Fuelpov2019) +
  geom_sf(aes(fill = pcFP)) +
  scale_fill_gradient(name = "Proportion of households fuel poor (%)", low="green",high="red")
```
### Difference in proportion Fuel Poverty 2012 to 2019
```{r Creating the idfference in proportion variable}
#Green indicates a smaller increase in fuel poverty from 2012 to 2019
#Red indicates an increase in fuel poverty from 2012 to 2019
fp_LA_2012_df <- readr::read_csv(paste0(dataFolder, "/Fuel Poverty/2012 LA_Fuel_Poverty.csv"))
fp_LA_2019_df <- readr::read_csv(paste0(dataFolder, "/Fuel Poverty/2019 LA_Fuel_Poverty.csv"))
merged_df <- merge(fp_LA_2012_df, fp_LA_2019_df, by = "Area Code", all = TRUE)
names(merged_df)
merged_df$pc_diff <-  as.numeric(merged_df$`Proportion of households fuel poor (%)`) - merged_df$`Proportion of households fuel poor (%) 2012`
```
```{r Mapping the iddference in fuel poverty from 2012 to 2019}
merged_df$lad18cd <- la_fuelpov_2019$`Area Code` # Creating a variable with the local authority code in the same name 
la_merged_Fuelpov20192012dif <- merge(la_sf_data,merged_df)# Merge the local Authority boundaries and the Fuel Poverty data
la_merged_Fuelpov20192012dif$pcFP <- as.numeric(la_merged_Fuelpov20192012dif$`Proportion of households fuel poor (%)`) # convert to numeric from char - why does it load as char??
ggplot2::ggplot(la_merged_Fuelpov20192012dif) +
  geom_sf(aes(fill = pc_diff)) +
  scale_fill_gradient(name = " Difference in proportion of households fuel poor (%)", low="green",high="red")
```
# Mapping the Middle Layer Super Output Areas (MSOA)
## Mapping Domestic
### Mapping the Electricity Consumption

```{r Loading in the MSOA electricity consumption data}
# electricity consumption data at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_DOM_ELEC_2019.csv")
msoa_elecData <- readr::read_csv(inFile)
head(msoa_elecData)
```

```{r Loading in the Boundary Data for the MSOA}
inf <-paste0(dataFolder, "boundaries/MSOA/msoa_solent.shp")
message("Loading Middle Layer Super Output Area (MSOA) from file")
#When using shp files you might also need to have other supplementary files in the same folder to allow it to run https://cfss.uchicago.edu/notes/simple-features/
msoa_sf_data <- sf::read_sf(inf) 
head (msoa_sf_data)
table(msoa_sf_data$LAD11NM)#how many MSOAs are in each LA
```

```{r Mapping the electrcity consumption of the MSOAs}
msoa_elecData$MSOA11CD <- msoa_elecData$ 'Middle Layer Super Output Area (MSOA) Code' #Creating a variable with the LA code and the name as in the sf_data
msoa_merged_sf <-merge(msoa_sf_data, msoa_elecData)
ggplot2::ggplot(msoa_merged_sf) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Electricity: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```
### Mapping the Gas
```{r Loading in the MSOA gas consumption data}
# gas consumption data at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_GAS_2019.csv")
msoa_gasData <- readr::read_csv(inFile)
head(msoa_gasData)
```
```{r Mapping the gas consumption of the MSOAs}
msoa_gasData$MSOA11CD <- msoa_gasData$ 'Middle Layer Super Output Area (MSOA) Code' 
msoa_merged_sf_domgas <-merge(msoa_sf_data, msoa_gasData)
ggplot2::ggplot(msoa_merged_sf_domgas) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```
> Does this match to non-gas areas??

####Mapping the change in gas consumption from 2010 to 2019
```{r Creating the dfference in consumption variable}
#Green indicates a decrease in gas consumption from 2010 to 2019
#Red indicates an increase in gas consumption from 2010 to 2019
domgas_MSOA_2010 <- readr::read_csv(paste0(dataFolder, "/energy/MSOA_GAS_2010.csv"))
domgas_MSOA_2019 <- readr::read_csv(paste0(dataFolder, "/energy/MSOA_GAS_2019.csv"))
domgas_MSOA_2010$domgas2010 <- domgas_MSOA_2010$`Mean consumption (kWh per meter)`
domgas_MSOA_2019$domgas2019 <- domgas_MSOA_2019$`Mean consumption (kWh per meter)`
merged_domgas <- merge(domgas_MSOA_2010, domgas_MSOA_2019, by = "Middle Layer Super Output Area (MSOA) Code", all = TRUE)
names(merged_domgas)
merged_domgas$mc_diff <-  as.numeric(merged_domgas$domgas2019) - merged_domgas$domgas2010
summary(merged_domgas)
```

```{r Mapping the change in dom mean gas comsumption}
merged_domgas$MSOA11CD <- merged_domgas$"MSOA Code" #creating a variable with the LSOA code in the same name as in sf_data
msoa_domgas20102019 <- merge(msoa_sf_data, merged_domgas) #merging these
ggplot2::ggplot(msoa_domgas20102019) +
  geom_sf(aes(fill=mc_diff))+ 
  scale_fill_continuous(name="Change in Gas: Mean kWh per meter", low="green",high="red")+
  labs(caption ="Solent (all MSOAs)")
##Error saying "Could not reallocate memory"
```

## Mapping Commercial
### Mapping Electricty
```{r Loading in the MSOA electricity consumption data for industry }
# electricity consumption data for industry at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_NONDOM_2019.csv")
msoa_elecDatanondom <- readr::read_csv(inFile)
head(msoa_elecDatanondom)
```

```{r Mapping the electrcity consumption of the MSOAs Industrial Sector}
msoa_elecDatanondom$MSOA11CD <- msoa_elecDatanondom$ 'Middle Layer Super Output Area (MSOA) Code' 
msoa_merged_sf_nondom <-merge(msoa_sf_data, msoa_elecDatanondom)
ggplot2::ggplot(msoa_merged_sf_nondom) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Electricity: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```
####


### Mapping Gas
```{r Loading in the MSOA gas consumption data for industry }
# gas consumption data for industry at MSOA level (pre downloaded)
inFile <-paste0(dataFolder, "energy/MSOA_NonDom_Gas_2019.csv")
msoa_gasDatanondom <- readr::read_csv(inFile)
head(msoa_gasDatanondom)
```

```{r Mapping the gas consumption of the MSOAs Industrial Sector}
msoa_gasDatanondom$MSOA11CD <- msoa_gasDatanondom$ 'Middle Layer Super Output Area (MSOA) Code' 
msoa_merged_sf_nondom_gas <-merge(msoa_sf_data, msoa_gasDatanondom)
ggplot2::ggplot(msoa_merged_sf_nondom_gas) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all MSOAs)")
#mapping it
```
> The holes tell you where there is no non-domestic gas use

####Mapping the difference in non domestic gas from 2010 to 2019 

here here here











# Mapping the Lower Layer Super Output Area (LSOA)

## Mapping Domestic

### Mapping Electricity consumption

```{r Loading in the LSOA electricty consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/LSOA_ELEC_2019.csv")
lsoa_elecData <- readr::read_csv(inFile)
lsoa_elecData$LSOA11CD <- lsoa_elecData$`Lower Layer Super Output Area (LSOA) Code`
head(msoa_elecData)
```

```{r Loading in boundary data for the LSOA}
#The LSOA boundaries for the Solent have been pre-downloaded
inf<-paste0(dataFolder, "boundaries/LSOA/lsoa_solent.shp")
message ("Loading LSOA boundaries from file")
lsoa_sf_data <- sf::read_sf(inf)
head(lsoa_sf_data)
table(lsoa_sf_data$LAD11NM)#How many LSOAs are there in each LA
```

```{r Mapping the electricty consumption of the LSOAs}
lsoa_merged_sf <- merge(lsoa_sf_data, lsoa_elecData) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf) + 
  geom_sf(aes(fill = `Mean domestic electricity consumption \n(kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all LSOAs)")
#Mapping it
```

#### Looking at cities in more detail
```{r Looking at the cities in more detail}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_merged_sf, LAD11NM == "Southampton")

# plotting the electricity consumption of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `Mean domestic electricity consumption \n(kWh per meter)`)) +
  scale_fill_continuous(name = "Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Southampton")
```


### Mapping Gas

```{r Loading in the LSOA gas consumption data}
#Electricty consumption data LSOA pre downloaded
inFile <-paste0(dataFolder, "energy/LSOA_GAS_2019.csv")
lsoa_gasData <- readr::read_csv(inFile)
lsoa_gasData$LSOA11CD <- lsoa_gasData$`Lower Layer Super Output Area (LSOA) Code`
head(lsoa_gasData)
```

```{r Mapping the gas consumption of the LSOAs}
lsoa_merged_sf_gas <- merge(lsoa_sf_data, lsoa_gasData) #merging the boundaries and energy data
ggplot2::ggplot(lsoa_merged_sf_gas) + 
  geom_sf(aes(fill = `Mean consumption (kWh per meter)`)) +
  scale_fill_continuous(name = "Gas: Mean kWh per meter", low = "green", high = "red") +
  labs(caption = "Solent (all LSOAs)")
#Mapping it
```

> Aha - this one shows where there is no gas :-)


### Mapping the Indices of Multiple Deprivation

```{r Loading in the Indices of Multiple Deprivation data}
# Multiple deprivation data at MSOA pre downloaded
inFile <-paste0(dataFolder, "Indices of Multiple Deprivation/Indices_of_Multiple_Deprivation_(IMD)_2019.csv")
lsoa_multidep <- readr::read_csv (inFile)
head(lsoa_multidep)
```

```{r Mapping the Indices of deprivation for the LSOA}
lsoa_multidep$LSOA11CD <- lsoa_multidep$lsoa11cd #creating a variable with the LSOA code in the same name as in sf_data
head(lsoa_multidep)
lsoa_multidep_merged_sf <- merge(lsoa_sf_data, lsoa_multidep) #merging these
ggplot2::ggplot(lsoa_multidep_merged_sf) +
  geom_sf(aes(fill=`IMDScore`))+ # which data set has the scores?
  scale_fill_continuous(name="IMD Score", low="green",high="red")+
  labs(caption ="Solent (all LSOAs)")
```
#####Looking at the cities in more detail
```{r Looking at Southampton in more detail}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_multidep_merged_sf, LAD11NM == "Southampton")

# plotting the multiple scales on dep of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `IMDScore`)) +
  scale_fill_continuous(name = "IMD Score", low = "green", high = "red") +
  labs(caption = "Southampton")
```
```{r Looking at Portsmouth in more detail}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_multidep_merged_sf, LAD11NM == "Portsmouth")

# plotting the change in fuel poverty from 2018-2012 of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `IMDScore`)) +
  scale_fill_continuous(name = "IMD Score", low = "green", high = "red") +
  labs(caption = "Portsmouth")
```
```{r Looking at Winchester in more detail}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_multidep_merged_sf, LAD11NM == "Winchester")

# plotting the fuel poverty 2019 of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = `IMDScore`)) +
  scale_fill_continuous(name = "IMD Score", low = "green", high = "red") +
  labs(caption = "Winchester")
```
### Mapping Fuel Poverty
#### 2019 Fuel Poverty
```{r Loading in the LSOA 2021 Fuel Poverty data}
inFile<-paste0(dataFolder, "Fuel Poverty/2019 LSOA_Fuel_Poverty.csv")
lsoa_fuelpov_2019<-readr::read_csv (inFile)
head(lsoa_fuelpov_2019)
```

```{r Mapping the LSOA 2019 Fuel Poverty}
lsoa_fuelpov_2019$LSOA11CD <- lsoa_fuelpov_2019$"LSOA Code" #creating a variable with the LSOA code in the same name as in sf_data
lsoa_fuelpov_2019_merged <- merge(lsoa_sf_data, lsoa_fuelpov_2019) #merging these
# as above - why does it load as a char?
lsoa_fuelpov_2019_merged$pcFP <- as.numeric(lsoa_fuelpov_2019_merged$`Proportion of households fuel poor (%)`)
ggplot2::ggplot(lsoa_fuelpov_2019_merged) +
  geom_sf(aes(fill=pcFP))+ 
  scale_fill_continuous(name="Proportion of households fuel poor (%)", low="green",high="red")+
  labs(caption ="Solent (all LSOAs)")
```
#####Looking at the cities in more detail
```{r Looking at Southampton in more detail}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov_2019_merged, LAD11NM == "Southampton")

# plotting the fuel poverty 2019 of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pcFP)) +
  scale_fill_continuous(name = "Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Southampton")
```
```{r Looking at Portsmouth in more detail}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov_2019_merged, LAD11NM == "Portsmouth")

# plotting the change in fuel poverty from 2018-2012 of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pcFP)) +
  scale_fill_continuous(name = "Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Portsmouth")
```
```{r Looking at Winchester in more detail}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov_2019_merged, LAD11NM == "Winchester")

# plotting the fuel poverty 2019 of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pcFP)) +
  scale_fill_continuous(name = "Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Winchester")
```

#### 2018 Fuel Poverty
```{r Loading in the LSOA 2018 Fuel Poverty data}
inFile<-paste0(dataFolder, "Fuel Poverty/2018 LSOA_Fuel_Poverty.csv")
lsoa_fuelpov_2018<-readr::read_csv (inFile)
head(lsoa_fuelpov_2018)
```

```{r Mapping the LSOA 2018 Fuel Poverty}
lsoa_fuelpov_2018$LSOA11CD <- lsoa_fuelpov_2018$"LSOA Code" #creating a variable with the LSOA code in the same name as in sf_data
lsoa_fuelpov_2018_merged <- merge(lsoa_sf_data, lsoa_fuelpov_2018) #merging these
# as above - why does it load as a char?
lsoa_fuelpov_2018_merged$pcFP18 <- as.numeric(lsoa_fuelpov_2018_merged$`Proportion of households fuel poor (%)`)
ggplot2::ggplot(lsoa_fuelpov_2018_merged) +
  geom_sf(aes(fill=pcFP18))+ 
  scale_fill_continuous(name="Proportion of households fuel poor (%)", low="green",high="red")+
  labs(caption ="Solent (all LSOAs) 2018")
```
#####Looking at the cities in more detail
```{r Looking at Southampton in more detail}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov_2018_merged, LAD11NM == "Southampton")

# plotting the fuel poverty 2019 of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pcFP18)) +
  scale_fill_continuous(name = "Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Southampton 2018")
```
```{r Looking at Portsmouth in more detail}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov_2018_merged, LAD11NM == "Portsmouth")

# plotting the change in fuel poverty from 2018-2012 of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pcFP18)) +
  scale_fill_continuous(name = "Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Portsmouth 2018")
```
```{r Looking at Winchester in more detail}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov_2018_merged, LAD11NM == "Winchester")

# plotting the fuel poverty 2019 of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pcFP18)) +
  scale_fill_continuous(name = "Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Winchester 2018")
```
####Mapping the difference in proportion of fuel poverty from 2012 to 2018

```{r Creating the dfference in proportion variable}
#Green indicates a smaller increase in fuel poverty from 2012 to 2018
#Had to use 2018 as the metric for fuel poverty has changed in 2019 
#Red indicates an increase in fuel poverty from 2012 to 2018
fp_LSOA_2012_df <- readr::read_csv(paste0(dataFolder, "/Fuel Poverty/2012 LSOA_Fuel_Poverty.csv"))
fp_LSOA_2018_df <- readr::read_csv(paste0(dataFolder, "/Fuel Poverty/2018 LSOA_Fuel_Poverty.csv"))
fp_LSOA_2012_df$pc2012 <- fp_LSOA_2012_df$`Proportion of households fuel poor (%)`
fp_LSOA_2018_df$pc2018 <- fp_LSOA_2018_df$`Proportion of households fuel poor (%)`
merged_df <- merge(fp_LSOA_2012_df, fp_LSOA_2018_df, by = "LSOA Code", all = TRUE)
names(merged_df)
merged_df$pc_diff <-  as.numeric(merged_df$pc2018) - merged_df$pc2012
summary(merged_df)
```
```{r Mapping the change in Fuel Poverty}
merged_df$LSOA11CD <- merged_df$"LSOA Code" #creating a variable with the LSOA code in the same name as in sf_data
lsoa_fuelpov20122018 <- merge(lsoa_sf_data, merged_df) #merging these
ggplot2::ggplot(lsoa_fuelpov20122018) +
  geom_sf(aes(fill=pc_diff))+ 
  scale_fill_continuous(name="Change in Proportions of households fuel poor (%)", low="green",high="red")+
  labs(caption ="Solent (all LSOAs)")
```
##### Looking at cities in more detail
```{r Looking at Southampton in more detail}
#Cities such as Southampton disappear due to their density, we will now map Southampton alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov20122018, LAD11NM == "Southampton")

# plotting the change in fuel poverty from 2018-2012 of Southampton in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pc_diff)) +
  scale_fill_continuous(name = "Change in Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Southampton")
#Around Bargate and Freemantle increased fuel poverty
```
```{r Looking at Portsmouth in more detail}
#Cities such as Portsmouth disappear due to their density, we will now map Portsmouth alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov20122018, LAD11NM == "Portsmouth")

# plotting the change in fuel poverty from 2018-2012 of Portsmouth in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pc_diff)) +
  scale_fill_continuous(name = "Change in Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Portsmouth")
#Charles Dickens increased fuel poverty
```
```{r Looking at Winchester in more detail}
#Cities such as Winchester disappear due to their density, we will now map Winchester alone to see this area in more detail
mapData <- dplyr::filter(lsoa_fuelpov20122018, LAD11NM == "Winchester")

# plotting the change in fuel poverty from 2018-2012 of Winchester in more detail
ggplot2::ggplot(mapData) + 
  geom_sf(aes(fill = pc_diff)) +
  scale_fill_continuous(name = "Change in Proportions of households fuel poor (%)", low = "green", high = "red") +
  labs(caption = "Winchester")
```
